Robust Control of a Fluxonium Qubit

### 2020-11-30 paper/7conclusion.tex
updated figures, lots of proofreading

%% S7
\section{Conclusion}
In conclusion, we have demonstrated techniques for achieving robustness to systematic
errors and mitigating decoherence on a quantum system using state-of-the-art trajectory
optimization. We have proposed a scheme for mitigating longitudinal relaxation with time-optimal
control and an efficient optimization metric that comes at a constant computational cost as
opposed to integrating a master equation which scales quadratically with
the dimension of the Hilbert space.
We have proposed the derivative method for robust control which achieves
super-linear gate error reductions in the gate duration for the problem we studied here.
We have shown that robust control techniques can be used to mitigate decoherence due
to 1/$f$ flux noise, a dominant source of coherent errors for flux controlled qubits.
The numerical techniques we have studied here can be used to perform phase gates in arbitrary times,
which will be critical for synchronizing multi-qubit systems. Our
techniques significantly improve the state-of-the-art for controlling quantum systems subject
to systematic errors and decoherence, and will be used in efforts to lower gate errors
for fault-tolerant quantum computing and NISQ applications.

>>>>

\section{Conclusion}
In conclusion, we have applied state-of-the-art trajectory
optimization techniques to achieve robustness to parameter deviations
and mitigate decoherence on a quantum system.
We have proposed a scheme for mitigating
longitudinal relaxation with time-optimal
control and the probability of longitudinal relaxation metric,
avoiding the optimal control problem for density matrices which scales quartically with
the dimension of the Hilbert space.
We have proposed the derivative method for robust control which achieves
super-linear gate error reductions in its gate time for the static parameter
deviation problem we studied.
We have shown that the derivative, sampling, and unscented sampling methods
can mitigate 1/$f$ flux noise errors
which dominate coherent errors for flux controlled qubits.
The derivative and sampling methods scale cubically with
the dimension of the Hilbert space, allowing for applications
to larger quantum systems.
The derivative, sampling, and unscented sampling
methods will benefit from interleaving optimization with
experimental characterization, while the sampling and unscented
sampling methods can also be applied to error channels without models.
These techniques will be used to achieve the low gate errors
required for fault-tolerant quantum computing applications. Our
implementation of the techniques described in this work is available
at \url{https://github.com/SchusterLab/rbqoc}.

### 2020-11-30 paper/a1longitude.tex
updated figures, lots of proofreading

\section{Longitudinal Relaxation}
\label{appendix:longitude}
We comment on the longitudinal relaxation metrics and then give
our procedure for integrating the Lindblad master equation.
The longitudal relaxation probability and the gate error due to longitudinal
relaxation metrics are compared in Table \ref{tab:longitude} for the experiment shown in Figure \ref{fig:longitude}c.
The relative performance of the analytic and numerical techniques
is similar across the two metrics.

\begin{table}[h!]
  \begin{tabular}{c | c | c | c | c | c | c}
         & $P_{1\textrm{A}}$ & $P_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
    Gate & $(10^{-5})$ & $(10^{-5})$ & $P_{1\textrm{A}}/ P_{1\textrm{N}}$ & $(10^{-5})$ &
    $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
    Y/2 & 5.253 & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
  \end{tabular}
  \caption{
    Single gate longitudinal relaxation probability ratios
    and single gate error due to longitudinal relaxation ratios.
    Values are reported for the analytic (A) and numerical (N) techniques
    presented in the main text.
  }
  \label{tab:longitude}
\end{table}

To compute the gate error due to longitudinal relaxation,
we require the final state of the quantum system subject
to longitudinal relaxation. To obtain the final state we employ
the Lindblad master equation. This equation takes the form
\begin{equation}
  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
  + \sum_{i = 1}^{n^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
\end{equation}
Here $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n = \textrm{dim}(\mathcal{H})$,
and $[\cdot, \cdot], \{\cdot, \cdot \}$ are the algebraic commutator and anti-commutator.
For longitudinal relaxation $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
$\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,
$L_{\uparrow} = \sigma^{+}/2$, and
$L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline shown in Figure \ref{fig:longitude}b.
The $T_{1}$ values in this spline are obtained experimentally by driving the qubit at the desired flux amplitude
and monitoring the resultant decay. For more details consult \cite{zhang2020universal}.

Exponential integrators can be employed to integrate the Lindblad master equation
using the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018}
\begin{equation}
  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho})
\end{equation}
\begin{equation}
  \begin{aligned}
    \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
    &+ \sum_{i = 1}^{n^{2} - 1} \gamma_{i}
    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
    - L_{i}^{T}L_{i}^{*} \otimes I))
  \end{aligned}
\end{equation}
Here $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
We use zero-order hold on the controls so the integration is exact
$\textrm{vec}(\rho_{k + 1}) = \exp(\Delta t_{k} \hat{\mathcal{L}}_{k}) \textrm{vec}(\rho_{k})$.
This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$
matrix-matrix multiplications to $(n^{4} \times n^{4}) \times n^{4}$ matrix-vector
multiplications. For small $n$ and zero-order held controls, we find that it is
faster to use an exponential integrator on the vectorized equation than to perform
Runge-Kutta on the unvectorized equation. The latter requires decreasing the integration
time step to maintain accuracy, resulting in more knot points.

>>>>

\section{Longitudinal Relaxation \label{appendix:longitude}}
We comment on the longitudinal relaxation metrics and then give
our procedure for integrating the Lindblad master equation.
The longitudinal relaxation probability and the gate error due to longitudinal
relaxation metrics are compared in Table \ref{tab:longitude} for the
numerical experiment described in Section \ref{sec:longitude}.
The relative performance of the analytic and numerical techniques
is similar across the two metrics.

\begin{table}[h!]
  \begin{tabular}{c | c | c | c | c | c | c}
         & $P_{1\textrm{A}}$ & $P_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
    Gate & $(10^{-5})$ & $(10^{-5})$ & $P_{1\textrm{A}}/ P_{1\textrm{N}}$ & $(10^{-5})$ &
    $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
    Y/2 & 5.253 & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
  \end{tabular}
  \caption{
    Single gate longitudinal relaxation probability (P) ratios
    and single gate error due to longitudinal relaxation (GE) ratios.
    Values are reported for the analytic (A) and numerical (N) techniques.
  }
  \label{tab:longitude}
\end{table}

To compute the gate error due to longitudinal relaxation
we employ the Lindblad master equation. This equation takes the form:
\begin{equation}
  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
  + \sum_{i = 1}^{n^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\})
\end{equation}
where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n$ is the
dimension of the Hilbert space,
$[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
anti-commutator.
For longitudinal relaxation $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
$\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,
$L_{\uparrow} = \sigma^{+}/2$, and
$L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline
shown in Figure \ref{fig:longitudeb}.
We obtain the $T_{1}$ values in this spline
by driving the qubit at the desired flux amplitude
and monitoring the resultant decay.
For more details consult \cite{zhang2020universal}.

Exponential integrators can be employed to integrate the Lindblad master equation
using the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
\begin{equation}
  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho})
\end{equation}
\begin{equation}
  \begin{aligned}
    \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
    &+ \sum_{i = 1}^{n^{2} - 1} \gamma_{i}
    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
    - L_{i}^{T}L_{i}^{*} \otimes I))
  \end{aligned}
\end{equation}
where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
We use zero-order hold on the controls so the exact solution is
$\textrm{vec}(\rho_{k + 1}) = \exp(\Delta t_{k} \hat{\mathcal{L}}_{k}) \textrm{vec}(\rho_{k})$.
This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$
matrix-matrix multiplications to $(n^{4} \times n^{4}) \times n^{4}$ matrix-vector
multiplications. For small $n$ and zero-order hold on the controls, we find that it is
faster to use an exponential integrator on the vectorized equation than to perform
Runge-Kutta on the unvectorized equation. The latter requires decreasing the integration
time step to maintain accuracy, resulting in more knot points.

### 2020-11-30 paper/a2unscented.tex
updated figures, lots of proofreading

\section{Unscented Transformation}
\label{appendix:unscented}
In this section we outline the full unscented sampling procedure.
The unscented sampling method concerns a quantum state $\psi \in \mathbb{R}^{2n}$ with
deviant parameters $\lambda \in \mathbb{R}^{d}$ and dynamics $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
The nominal initial quantum state is given by $\psi_{0}$ with an associated
positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{n \times n}$
which describes the uncertainty in the initial state.
$P_{0}$ is typically chosen to be non-zero even if the state preparation error is negligible.
The deviant parameter
is assumed to have zero mean and its distribution is given by the covariance matrix
$L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
is convenient for deriving the update procedure. A non-zero mean can be encoded
in the dynamics $f$.

The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
from the initial distributions
\begin{equation}\label{eq:uupdate}
  \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
  \begin{bmatrix} \bar{\Psi}_{0} \\ 0\end{bmatrix}
    \pm \beta \sqrt{\begin{bmatrix} P_{0} & 0\\ 0 & L_{0}\end{bmatrix}}^{\; i}
\end{equation}
We have written $\bar{\Psi}_{0} = \psi_{0}$. $\beta$
is a hyperparameter which controls the spacing of the covariance contour.
The $(\pm)$ is understood to take $(+)$ for $ i \in \{1, \dots, 2n + d\}$ and $(-)$ for
$i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
to compute the square root of the
joint covariance matrix, though other methods
such as the principal square root may be employed.
The superscript on the matrix square root indicates the $i$\textsuperscript{th}
column (mod $2n + d$) of the lower triangular Cholesky factor.
Then, the sigma points are normalized
\begin{equation}\label{eq:unormalize}
  \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
\end{equation}
The sigma points are propagated to the next knot point
\begin{equation}\label{eq:upropagate}
  \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
\end{equation}
The mean and covariance of the sigma points are computed
\begin{align}
  \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
  P_{1} &= \frac{1}{2 \beta^{2}} \sum_{i = 1}^{4n + 2d}
  (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
\end{align}
The sigma points are then resampled and propagated to the next knot point using
\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
choice of sigma points follows the prescription in equation 11 of \cite{julier2004unscented}.
Other prescriptions may be chosen to reduce the number of required sigma points \cite{julier2002reduced}.

>>>>

\section{Unscented Transformation}
\label{appendix:unscented}
In this section, we outline the full unscented sampling procedure.
We consider a state $\psi \in \mathbb{R}^{2n}$ with
a deviant parameter $\lambda \in \mathbb{R}^{d}$ and dynamics
$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
The nominal initial state is given by $\psi_{0}$ with an associated
positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{2n \times 2n}$
which describes the uncertainty in the initial state.
$P_{0}$ is typically non-zero even if the state preparation error is negligible.
The deviant parameter
has zero-mean and its distribution is given by the covariance matrix
$L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
is convenient for deriving the update procedure. A non-zero mean can be encoded
in the dynamics.

The initial $4n + 2d$ sigma points and initial $4n + 2d$ deviant parameters are sampled
from the initial distributions,
\begin{equation}\label{eq:uupdate}
  \begin{bmatrix} \Psi_{0}^{i} \\ \Lambda_{0}^{i} \end{bmatrix} =
  \begin{bmatrix} \bar{\Psi}_{0} \\ 0\end{bmatrix}
    \pm \beta \sqrt{\begin{bmatrix} P_{0} & 0\\ 0 & L_{0}\end{bmatrix}}^{\; i}
\end{equation}
We have written $\bar{\Psi}_{0} = \psi_{0}$. $\beta$
is a hyperparameter which controls the spacing of the covariance contour.
The $(\pm)$ is understood to take $(+)$ for $ i \in \{1, \dots, 2n + d\}$ and $(-)$ for
$i \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
to compute the square root of the
joint covariance matrix, though other methods
such as the principal square root may be employed.
The superscript on the matrix square root indicates the $i$\textsuperscript{th}
column (mod $2n + d$) of the lower triangular Cholesky factor.
Then, the sigma points are normalized,
\begin{equation}\label{eq:unormalize}
  \Psi_{0}^{i} \gets \frac{\Psi_{0}^{i}}{\sqrt{\Psi_{0}^{i^{T}} \Psi_{0}^{i}}}
\end{equation}
The sigma points are propagated to the next knot point,
\begin{equation}\label{eq:upropagate}
  \Psi_{1} = f(\Psi_{0}, \Lambda_{0})
\end{equation}
The mean and covariance of the sigma points are computed,
\begin{align}
  \bar{\Psi}_{1} &= \frac{1}{4n + 2d} \sum_{i = 1}^{4n + 2d} \Psi_{1}^{i}\\
  P_{1} &= \frac{1}{2 \beta^{2}} \sum_{i = 1}^{4n + 2d}
  (\Psi^{i}_{1} - \bar{\Psi}_{1})^{T}(\Psi^{i}_{1} - \bar{\Psi}_{1})
\end{align}
The sigma points are then resampled and propagated to the next knot point using
\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
choice of sigma points follows the prescription in equation 11 of \cite{julier2004unscented}.
Prescriptions that require fewer sigma points exist \cite{julier2002reduced}.

### 2020-11-30 paper/a3derivative.tex
updated figures, lots of proofreading

\section{Derivative Method \label{appendix:derivative}}
We comment on the optimization metrics of the derivative methods and then
outline how to efficiently integrate their dynamics.
The 1\textsuperscript{st}-order derivative method tends
to outperform the 2\textsuperscript{nd}-order
derivative method in the small, static detuning regime, see Figure \ref{fig:static}b.
The norms of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
methods are provided in Table \ref{tab:dnorm} for the solutions at $t_{N} = 60$ns.
The 2\textsuperscript{nd}-order method is able to decrease the 2\textsuperscript{nd}-order
state derivate norm relative to the 1\textsuperscript{st}-order method at the expense of increasing
its 1\textsuperscript{st}-order derivative norm. In the problem
we have studied it is likely the 2\textsuperscript{nd}-order derivative norm has a smaller
contribution to the gate error than the 1\textsuperscript{st}-order derivative norm. A
careful analysis could be completed for future problems to predict the efficacy of the
derivative method at each order.

\begin{table}[ht]
  \begin{tabular}{c | c | c}
    Method & ${\lvert \braket{\partial_{f_{q}} \psi_{N} | \partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
    & ${\lvert \braket{\partial^{2}_{f_{q}} \psi_{N} | \partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{6}$)\\
    \hline
    D-1 & 0.436 & 57.817\\
    D-2 & 1.702 & 9.030\\
  \end{tabular}
  \caption{Norm of state derivates with respect to the qubit frequency
    for $Z/2$ gates optimized using the derivative methods. The norms are computed
    at the end of the gate's duration $t_{N} = 60$ns and are averaged over the four state derivatives.}
  \label{tab:dnorm}
\end{table}

The dynamics for the derivative methods can be integrated efficiently using exponential integrators
\cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
General exponential integrators break the dynamics into a linear term and a non-linear term.
For example, consider
integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
With zero-order hold on the controls the exact propagation is
\begin{equation}
  \label{eq:dgeneralexp}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k}) N(t_{k} + t^{'}) dt^{'}\\
  \end{aligned}
\end{equation}
General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
into a discrete sum, similar to the procedure
for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
method \cite{berland2005solving}
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
  \end{aligned}
\end{equation}
This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
exponential computation per stage. Integration accuracy is not of the utmost importance because the
state derivatives guide the optimization, and do not correspond to experimental parameters
which must be realized with high accuracy.

>>>>

\section{Derivative Method \label{appendix:derivative}}
We comment on the optimization metrics of the derivative methods and then
outline how to efficiently integrate their dynamics.
The 1\textsuperscript{st}-order derivative method tends
to outperform the 2\textsuperscript{nd}-order
derivative method in the small, static detuning regime, see Figure \ref{fig:staticb}.
The norms of the state derivatives for the 1\textsuperscript{st}- and 2\textsuperscript{nd}-order
methods are provided in Table \ref{tab:dnorm} for the solutions to
the numerical experiment detailed in Section \ref{sec:static} at $t_{N} = 60$ns.
The 2\textsuperscript{nd}-order method is able to decrease the 2\textsuperscript{nd}-order
state derivate norm relative to the 1\textsuperscript{st}-order method at the expense of increasing
its 1\textsuperscript{st}-order derivative norm. In the problem
we have studied it is likely the 2\textsuperscript{nd}-order derivative norm has a smaller
contribution to the gate error than the 1\textsuperscript{st}-order derivative norm. A
careful analysis of contributions at each order could be used to predict the efficacy of the
derivative method for future problems.

\begin{table}[ht]
  \begin{tabular}{c | c | c}
    Method & $\norm{\partial_{f_{q}} \psi_{N}}_{2}$ ($10^{3}$)
    & $\norm*{\partial^{2}_{f_{q}} \psi_{N}}_{2}$ ($10^{6}$)\\
    \hline
    D1 & 0.436 & 57.817\\
    D2 & 1.702 & 9.030\\
  \end{tabular}
  \caption{Norm of state derivates with respect to the qubit frequency
    for $Z/2$ gates optimized using the derivative methods. The norms are computed
    at the end of the gate $t_{N} = 60$ns
    and are averaged over the four state derivatives.}
  \label{tab:dnorm}
\end{table}

The dynamics for the derivative methods can be integrated efficiently using exponential integrators
\cite{auer2018magnus, berland2005solving, einkemmer2017performance}.
General exponential integrators break the dynamics into a linear term and a non-linear term.
For example, consider
integrating the dynamics of the first state derivative $\frac{d}{dt} \ket{\partial_{\lambda} \psi} =
H \ket{\partial_{\lambda} \psi} + (\partial_{\lambda} H) \ket{\psi}$ in units of $i\hbar = 1$.
The linear term is $L = H$ and the non-linear term is $N = (\partial_{\lambda} H) \ket{\psi}$.
With zero-order hold on the controls the exact solution is:
\begin{equation}
  \label{eq:dgeneralexp}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &= \exp(\Delta t_{k} L_{k})
    \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \int_{0}^{\Delta t_{k}} \exp((\Delta t_{k} - t^{'})L_{k})
    N(t_{k} + t^{'}) dt^{'}\\
  \end{aligned}
\end{equation}
General exponential integrators proceed by breaking the integral in \eqref{eq:dgeneralexp}
into a discrete sum, similar to the procedure
for Runge-Kutta schemes. We use a simple approximation known as the Lawson-Euler
method \cite{berland2005solving},
\begin{equation}
  \begin{aligned}
    \ket{\partial_{\lambda} \psi_{k + 1}} &\approx \exp(\Delta t_{k}L_{k}) \ket{\partial_{\lambda} \psi_{k}}\\
    &+ \exp(\Delta t_{k}L_{k}) N_{k}\\
  \end{aligned}
\end{equation}
This method provides a good tradeoff between accuracy and efficiency, requiring one unique matrix
exponential computation per stage. Integration accuracy is not of the utmost importance because the
state derivatives guide the optimization, and do not correspond to experimental parameters
which must be realized with high accuracy.

### 2020-11-30 paper/a4time.tex
prog
updated figures, lots of proofreading

\section{Computational Performance and Scaling}
\todo{timing table}

\todo{make complexity analysis a coherent paragraph}
(sampling) For the sampling method the number of sample states in the augmented state vector
scales as $O(dn^{2})$
because each of $n^{2}$ initial
states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
and the size of a state vector. $d$ is the number of deviant parameters.

(unscented) Each initial state from the operator basis
is represented with a distribution of $4n + 2d$ samples. Hence,
the number of states in the augmented state vector scales as $O(n^{3} + dn^{2})$.
The gate error of each sample state is penalized as in the sampling method.

(derivative) For this method $m$ state derivatives are associated with each
initial state from the operator basis.
So, the number of states in the augmented state vector scales as
$O(dmn^{2})$. Additionally, we note that the number of initial
states and samples we have employed for each of the three
methods follows standard perscriptions. However, knowledge
of the problem can typically be used to reduced the number
of required samples.

>>>>

\section{Computational Performance and Scaling \label{appendix:time}}
We provide runtimes for our optimizations and comment on the scaling of the
robustness methods. The runtimes for the base optimization in Section \ref{sec:fluxonium},
longitudinally aware optimization in Section \ref{sec:longitude},
and robust optimizations in Section \ref{sec:static} are presented
in Table \ref{tab:time} for a $Z/2$ gate at multiples
of the analytic gate time.
\todo{timing table}

\begin{table}[ht]
  \begin{tabular} {c | c | c | c | c | c | c}
    & \multicolumn{6}{c}{Wall Time ($\pm 0.001$ s)}\\
    \hline
    Gate Time (ns) & Base & Long. & S & SU & D1 & D2\\
    \hline
    $18$ & 0.155 & 1.688 & 1.773 & & 16.713 & 48.398\\ % 0.155049
    $36$ & 7.014 & - & 48.213 & & 67.838 & 81.030\\
    $72$ & 15.906 & - & & & &\\
  \end{tabular}
  \caption{Programs were executed on a single thread of an AMD Ryzen Threadripper 3970X 32-Core Processor.}
  \label{tab:time}
\end{table}

\todo{Ask Dave if we should keep complexity analysis in main text.}
Now we present the problem size complexities for the robustness methods.
For the sampling method, the size of the augmented state vector
is $O(dn^{3})$. $d$ is the number of deviant parameters.
$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
in the operator basis, $2d$ sample states per initial state,
and each state has $2n$ elements.
For the unscented sampling method, the size of the augmented state vector
is $O(dn^{3} + n^{4})$.
There are $n^{2}$ initial states in the operator basis,
$2(2n + d)$ sample states per initial state,
and each state has $2n$ elements.
For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
is $O(dmn^{3})$. There are $n^{2}$
initial states in the operator basis, $dm$ state derivatives per initial state,
and each state has $2n$ elements.
